---
title: "tidy coral with Jeanette"
author: "Mary K. Donovan"
date: "3/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r}
## libraries
library(tidyverse)
library(janitor) # install.packages('janitor')
library(skimr) # install.packages('skimr')
library(stringr) # added when we needed it for benthic data

## data filepaths/urls ----

## benthic data
benthic_url <- 'https://www.nodc.noaa.gov/archive/arc0054/0104255/1.1/data/0-data/cd08/100308OaAla03m.CSV'

## buoy data
buoy_url <- 'http://www.ndbc.noaa.gov/view_text_file.php?filename=mokh1h2010.txt.gz&dir=data/historical/stdmet/'
```

## Data
This is benthic data from a series of CRAMP (Coral Reef Assessment Monitoring Program) data that includes Kaneohe Bay coral survey still images and extracted data (with larger Hawaiian Islands context):

```{r}
benthic.raw <- read_csv(benthic_url)
head(benthic.raw)
```
*ignoring black values for now

## Wrangle
Janitor package will clean up the column names
```{r}
benthic <- benthic.raw %>% 
  janitor::clean_names()
names(benthic)
```

Some columns aren't useful
```{r}
benthic <- benthic %>% 
  select(id_name, point, x, y, id_date)
head(benthic)
```

Date column needs formatting
```{r}
benthic <- benthic %>% 
  mutate(date = stringr::str_remove_all(id_date,"#"))
skimr::skim(benthic)
```

Plot
```{r}
ggplot(benthic, aes(x=date, fill=id_name)) + geom_bar()
```

## Buoy data

read in data
```{r}
buoy <- readr::read_csv(buoy_url)
head(buoy)
```
Not the correct delim

```{r}
buoy <- readr::read_table(buoy_url)
head(buoy)
```
Nope still not right (cols missing)

```{r}
buoy_raw <- readr::read_table2(buoy_url)
head(buoy_raw)
buoy <- buoy_raw
```

### Wrangle
The buoy data come from the National Buoy Data Center. We are going to use data from the inner Kaneohe Bay buoy (station MOKH1). More details on this station are available here.

There’s 3 things we want to do to these column names:
1. make the column header a combo of rows 1 and 2
```{r}
names(buoy) <- str_c(names(buoy),                  ## current header
                  buoy %>% filter(`#YY` == "#yr"), ## row1 -- don't say buoy[1,] 
                  sep = "_")                       ## separate by `_`

## inspect
names(buoy) ## Looks a lot better
```

2. clean up the header; get rid of # and /
```{r}
names(buoy) <- str_replace_all(names(buoy), "#", "")  # replace `#` with nothing 
names(buoy) <- str_replace_all(names(buoy), "/", "_") # replace `/` with `_`

## inspect to make sure it worked
names(buoy)
```

3. delete the now-redundant row 1
```{r}
buoy <- buoy %>%
  filter(YY_yr != "#yr")
```

### Plot
```{r}
ggplot(buoy, aes(WTMP_degC)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=90))
```
Missing values are set to 999
```{r}
buoy$WTMP_degC[buoy$WTMP_degC=='999.0'] <- NA
summary(as.numeric(buoy$WTMP_degC))
```

Need to coerce all the columns into integers
```{r}

```

We want to join the buoy data to the benthic data using the date, so need to format dates
```{r}
buoy <- buoy %>%
  unite(date,c(YY_yr,MM_mo,DD_dy),sep="-")
head(buoy)
```

## Join

```{r}
bb_join <- left_join(benthic,buoy,by='date')
```
Didn't work because there are multiple temp records per day

Create a summary of temp
```{r}
temp.mean <- buoy %>% 
  group_by(date) %>% 
  mutate(WTMP_degC.n = as.numeric(WTMP_degC)) %>% 
  summarise(meantemp = mean(WTMP_degC.n,na.rm=T)) %>% 
  ungroup()
head(temp.mean)
```

Try to join them again
```{r}
bb_join <- left_join(benthic,temp.mean,by='date')
skimr::skim(bb_join)
```

Plot
```{r}
plot(as.Date(bb_join$date, '%Y-%m-%d'),bb_join$meantemp,type="l")
```

